\bf Visual Diagnostics of Dynamic Social Network Models

### 2018-11-20 paper-child.Rnw
revise abstract

\par The second method for net-vis uses the adjacency matrix of the network. The adjaceny matrix of a network, $\mathbf{A}$ is defined as:
\begin{equation}\label{eq:adjmat}
  A_{ij} =
  \begin{cases}
                                   1 &  \text{if an edge exists } i\to j, \quad i \neq j \\
                                   0 & \text{otherwise}
  \end{cases}
\end{equation}
%In this paper, we only consider the presence or absence of an edge, but if the network has weighted edges, $A_{ij}$ is the weight of the edge from $i \to j$. 
The adjacency matrix visualization for our toy example is also shown in Figure~\ref{fig:toyplots}. This visualization shows the network as a grid, with each row representing the ``from" node and each column representing the ``to" node of an edge. The grid space for edge $i\rightarrow j$ is filled in if $A_{ij} = 1$ and is empty if $A_{ij} = 0$. For undirected networks, $\mathbf{A}$ and the grid visualization are symmetric, with each edge represented twice.

@@

The two pieces of the model hierarchy each contribute to the entries of the intensity matrix to describe the rate of change between two network states. The entries of $\mathbf{Q}$ are defined as follows: let $b \neq c \in \{1, 2, \dots, 2^{n(n-1)} \}$ be indices of two different possible states of the network, $x^b, x^c \in \mathcal{X}$.  Then the $bc^{th}$ entry of $Q$ is:
\[ q_{bc} = \begin{cases}
     q_{ij}= \alpha_m p_{ij}(\boldsymbol{\beta}, x^b) & \text{if } \sum_{i,j} |x_{ij}^c - x_{ij}^b| = 1\\
      -\sum_{i\neq j} q_{ij} & \text{if } x^b = x^c
      0 & \text{otherwise} \\
   \end{cases}
\]
The rate of change between any two states, $x^b$ and $x^c$, that differ by only one tie $x_{ij}$, is the product of the rate at which actor $i$ gets to change a tie and the probability that the tie that will change is the tie to node $j$. The intensity matrix $\mathbf{Q}$ fully defines the model for network change, and is the foundation required for parameter estimation. %Estimating the parameters in these models is difficult, but thanks to the SIENA software \citep{RSiena}, we have an accessible way to fit SAOMs to network observations.

>>>>

\par The second method for net-vis uses the adjacency matrix of the network. The adjaceny matrix of a network, $\mathbf{A}$ is defined as:
\begin{equation}\label{eq:adjmat}
  A_{ij} =
  \begin{cases}
             1 &  \quad \text{if an edge exists } i\to j, \quad i \neq j \\
             0 & \quad \text{otherwise}
  \end{cases}
\end{equation}
%In this paper, we only consider the presence or absence of an edge, but if the network has weighted edges, $A_{ij}$ is the weight of the edge from $i \to j$. 
The adjacency matrix visualization for our toy example is also shown in Figure~\ref{fig:toyplots}. This visualization shows the network as a grid, with each row representing the ``from" node and each column representing the ``to" node of an edge. The grid space for edge $i\rightarrow j$ is filled in if $A_{ij} = 1$ and is empty if $A_{ij} = 0$. For undirected networks, $\mathbf{A}$ and the grid visualization are symmetric, with each edge represented twice.

@@

The two pieces of the model hierarchy each contribute to the entries of the intensity matrix to describe the rate of change between two network states. The entries of $\mathbf{Q}$ are defined as follows: let $b \neq c \in \{1, 2, \dots, 2^{n(n-1)} \}$ be indices of two different possible states of the network, $x^b, x^c \in \mathcal{X}$.  Then the $bc^{th}$ entry of $Q$ is:
\[ q_{bc} = \begin{cases}
     q_{ij} = \alpha_m p_{ij}(\boldsymbol{\beta}, x^b) & \text{ if } \sum_{i,j} |x_{ij}^c - x_{ij}^b| = 1 \\
      -\sum_{i\neq j} q_{ij} & \text{ if } x^b = x^c \\
      0 & \text{ otherwise} 
   \end{cases}
\]
The rate of change between any two states, $x^b$ and $x^c$, that differ by only one tie $x_{ij}$, is the product of the rate at which actor $i$ gets to change a tie and the probability that the tie that will change is the tie to node $j$. The intensity matrix $\mathbf{Q}$ fully defines the model for network change, and is the foundation required for parameter estimation. %Estimating the parameters in these models is difficult, but thanks to the SIENA software \citep{RSiena}, we have an accessible way to fit SAOMs to network observations.

### 2018-11-20 paper.Rnw
revise abstract

\documentclass[12pt]{article}
% \usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
% \geometry{letterpaper}                   % ... or a4paper or a5paper or ...
%\usepackage{graphicx}
\usepackage[font=small,skip=5pt]{caption}
\usepackage{subcaption}
\usepackage{animate}
\usepackage{afterpage}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amsfonts}
% \usepackage{color}
\usepackage{multirow}
\usepackage{rotating}
\usepackage[dvipsnames,svgnames,table]{xcolor}
\usepackage{hyperref}
\graphicspath{{figure/}}
% \usepackage{endfloat} % Figures to the end of the document

@@

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

@@

\if0\blind
{
  \title{\bf Visualizing Dynamic Social Network Models}
  \author{Sam Tyner \hspace{.2cm}\\
    Department of Statistics and Statistical Laboratory, Iowa State University\\
    and \\
    Heike Hofmann\\
    Department of Statistics and Statistical Laboratory, Iowa State University}
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Visualizing Dynamic Social Network Models}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
Stochastic actor-oriented models (SAOMs), first introduced by Snijders (1996), are statistical network models for dynamic social networks. Unlike other network models, SAOMs are analytically intractable and require extensive simulation procedures for model fitting. In this paper, we use model visualization techniques introduced in Wickham et al (2015) to better understand SAOMs and their behavior. We use a three-pronged visualization attack: (a) network visualizations to view the \emph{model}, (b) visualizations of collections of models to interpret parameters, and (c) visualizations of fitting algorithms to gain insights into the underlying mechanisms. With the help of static and dynamic visualizations, we bring the hidden SAOM fitting processes into the foreground, with the goal of a better understanding and higher accessibility of SAOMs by social network analysts.
\end{abstract}

>>>>

\documentclass[12pt]{article}
% \usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
% \geometry{letterpaper}                   % ... or a4paper or a5paper or ...
%\usepackage{graphicx}
\usepackage[font=small,skip=5pt]{caption}
\usepackage{subcaption}
\usepackage{animate}
\usepackage{afterpage}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amsfonts}
% \usepackage{color}
\usepackage{multirow}
\usepackage{rotating}
\usepackage[dvipsnames,svgnames,table]{xcolor}
\usepackage{hyperref}
\graphicspath{{figure/}}
%\usepackage{endfloat} % Figures to the end of the document

@@

\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

@@

\if0\blind
{
  \title{\bf Visual Diagnostics of Dynamic Social Network Models}
  \author{Sam Tyner \hspace{.2cm}\\
    Department of Statistics and Statistical Laboratory, Iowa State University\\
    and \\
    Heike Hofmann\\
    Department of Statistics and Statistical Laboratory, Iowa State University}
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Visual Diagnostics of Dynamic Social Network Models}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
Statistical models for networks have been studied for decades, and they have gotten more complex as computing power has grown. Some network models, such as those for dynamic social networks networks, are analytically intractable and therefore require extensive simulation procedures for model fitting. In this paper, we use model visualization techniques introduced in Wickham et al (2015) to better understand some of these models and their behavior. We use a three-pronged visualization attack: (a) network visualizations to view the \emph{model}, (b) visualizations of collections of models to interpret parameters, and (c) visualizations of fitting algorithms to gain insights into the underlying mechanisms. With the help of static and dynamic visualizations, we bring the hidden algorithmic fitting processes into the foreground, with the goal of a better understanding and higher accessibility of statistical models for social network analysts.
\end{abstract}
